\documentclass[sigconf]{acmart}
\settopmatter{printacmref=false} % Removes citation information below abstract
\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference information in first column
\pagestyle{plain} % removes running headers

\usepackage{booktabs} % For formal tables
\usepackage{graphicx} % Required for including images
\graphicspath{{figures/}} % Location of the graphics files
\usepackage{booktabs} % Top and bottom rules for table
\usepackage[font=small,labelfont=bf]{caption} % Required for specifying captions to tables and figures
\usepackage{amsfonts, amsmath, amsthm, amssymb} % For math fonts, symbols and environments
\usepackage{wrapfig} % Allows wrapping text around tables and figures
\usepackage{hyperref}
\usepackage{multirow}
\usepackage{rotating}
% \usepackage{subfig}% http://ctan.org/pkg/subfig
\usepackage {tikz}
\usetikzlibrary {positioning}
%\usepackage {xcolor}
\definecolor {processblue}{cmyk}{0.96,0,0,0}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{dblfloatfix}

% \makeatletter
% \def\@copyrightspace{\relax}
% \makeatother

% correct bad hyphenation here
\hyphenation{Page-Rank}

\begin{document}
\title{Information Retrieval on the UCL Website}
\subtitle{COMPGI15}


\author{Minttu Alakuijala}
\affiliation{\institution{}}
\email{minttu.alakuijala.14@ucl.ac.uk}

\author{Nathalie Von Huth}
\affiliation{\institution{}}
\email{nathalie.huth.14@ucl.ac.uk}

\author{Aoi Yamamoto}
\affiliation{\institution{}}
\email{aoi.yamamoto.16@ucl.ac.uk}

\author{Yuechen Zhang}
\affiliation{\institution{}}
\email{yuechen.zhang.16@ucl.ac.uk}

% The default list of authors is too long for headers}
% \renewcommand{\shortauthors}{B. Trovato et al.}


\begin{abstract}

Finding information on the World Wide Web has become of central importance to everyday affairs. Users often have certain needs such as relevancy, speed and accuracy of results. Over the past few decades, ranking algorithms in information retrieval systems have largely shaped not only how we surf the web, but also how we navigate private information such as emails and files in cloud storage. The goal of retrieval systems is to cover the user's needs by providing the most relevant documents given one or multiple search terms. The results need to be fair, justified, and able to capture the intent of the user. This report explores some of the most widely adopted algorithms and measures their performance against an original test set. Based on an index of the UCL Computer Science department domain (cs.ucl.ac.uk), we report results on Boolean retrieval, Term Frequency -- Inverse Document Frequency, BM25 and query likelihood ranking algorithms, each with the added option of including PageRank authority scores. After evaluating our results through precision, recall, F1 and discounted cumulative gain, we concluded that BM25 performs best out of the four ranking algorithms with scores of up to 70\% precision and up to 62\% recall, and managed to outperform the UCL search engine. TF-IDF and query likelihood follow closely behind BM25. It was also observed that for our restricted domain with relatively few internal hyperlinks, PageRank does not add relevance value, as it overemphasises pages with incoming hyperlinks.

\end{abstract}

\maketitle

\input{report-body}

\bibliographystyle{ACM-Reference-Format}
\bibliography{bibliography} 

\end{document}
